Analysis of the 2005--2015 Earthquake Sequence in Northern Iran Using the Visibility Graph Method

### 2016-08-22 abstract.tex
fixed minor details in text

%
We present an analysis of the seismicity of northern Iran in the period between 2005 and 2015 using a recently introduced method based on concepts of graph theory. The method relies on the inter-event visibility defined in terms of a connectivity degree parameter, $k$, which is correlated with the earthquake magnitude, $M$. Previous studies show that the slope of the $k$-$M$ correlation also observes a relationship with the $b$-value from the Gutenber-Richter law. Thus rendering the graph analysis useful to examine the seismicity of a region. In addition, other topological properties of the network and dynamic properties of the magnitude-time series exhibit characteristics that seem to be associated with the occurrence of earthquakes, offering the possibility of correlating seismicity parameters with time. We apply this approach to the case of the seismicity of northern Iran, using a declustered seismic catalogs for the tectonic seismic regions of Azerbaijan, Alborz, and Kopeh Dagh. We use results draw for this region with the visibility graph approach in combination with results from other similar studies to further improve the universal correlation between $k$-$M$ and $b$-values and show that the visibility graph approach can be considered as a valid alternative for analyzing regional seismicity properties and earthquake sequences.

% --------------------------------------------------------------------------------------------------

% NAEEM'S ABSTRACT
% ----------------

% The seismicity of northern Iran between 2005 and 2015 is investigated by means of the visibility graph (VG) method. For purposes of this study, northern Iran is divided into three tectonic seismic regions: Azerbaijan, Alborz, and Kopeh Dagh. Using the declustered catalog, we studied the VG properties of the magnitude time series. The results show a relationship between the $k-M$ slope and the $b-value$ of the Gutenberg-Richter law. Topological properties (i.e.  $< T_c >$, $k-M$ slope) of the network and dynamic properties of magnitude time series (i.e. $b-value$) significantly decrease before large earthquakes. Combining the results of this study with three similar previous studies improves the linear correlation coefficient factor between $k-M$ slope and $b-value$, and empowers the idea of a universal relationship between $b-value$ and $k-M$ slope. The behaviors of the VG's properties are similar to previous studies and could be considered an alternative method for analyzing earthquake sequences.

>>>>

%
We present an analysis of the seismicity of northern Iran in the period between 2005 and 2015 using a recently introduced method based on concepts of graph theory. The method relies on the inter-event visibility defined in terms of a connectivity degree parameter, $k$, which is correlated with the earthquake magnitude, $M$. Previous studies show that the slope of the $k$-$M$ correlation also observes a relationship with the $b$-value from the Gutenber-Richter law, thus rendering the graph analysis useful to examine the seismicity of a region. In addition, other topological properties of the network and dynamic properties of the magnitude-time series exhibit characteristics that seem to be associated with the occurrence of earthquakes, offering the possibility of correlating seismicity parameters with time. We apply this approach to the case of the seismicity of northern Iran, using an earthquake catalog for the tectonic seismic regions of Azerbaijan, Alborz, and Kopeh Dagh. We use results draw for this region with the visibility graph approach in combination with results from other similar studies to further improve the universal correlation between $k$-$M$ and $b$-values and show that the visibility graph approach can be considered as a valid alternative for analyzing regional seismicity properties and earthquake sequences.

### 2016-08-22 introduction.tex
fixed minor details in text

For the particular case of seismological applications, previous studies have proposed the nodes in the graph to be considered as the earthquakes in a given time series of events, and analyzed the properties of the graph with respect to traditional seismicity parameters. The analysis of seismic sequences through the visibility graph approach for various tectonic seismic regions has proven to be a valid alternative to studying magnitude time series. \citet{Telesca2012}, for instance, studied the seismicity of Italy between the years 2005 and 2010 using the visibility graph method. Applying different threshold magnitudes to construct the graph and observing the collapsing effect of all distribution degrees, they observed that the properties of the visibility graph seemed to depend only on the magnitude values, and not on the threshold magnitude used in the analysis.

@@

In a similar article, \citet{Telesca2014} studied the seismicity of the 2002--2011 Pannonian seismicity using the visibility graph method for two sub-catalogs of shallow and deep earthquakes. They extracted the visibility graph characteristics for each group of events and confirmed that there was a close correlation between the Gutenberg-Richter $b$-value and the slop of the $k$-$M$ relationship obtained from the visibility graph. According to \citet{Telesca2014}, the high linear correlation coefficient value (close to 1.0) between the $b$-value and the $k$-$M$ slope indicates that this relationship exhibits a nearly universal character. This observation was reinforced experimentally by \citet{Telesca2014-pone}, who investigated the behavior of a mechanical stick-slip system with different asperities (sandpaper of different grades). After collecting data for different experiments emulating young and mature faults, \citet{Telesca2014-pone} found again that the $b$-value was linearly correlated with the $k$-$M$ slope derived from the visibility graph of the experimental events.

More recently, \citet{Telesca2016} used the visibility graph method to gain insight about the timespan between earthquakes. The authors defined the parameter window mean interval connectivity time <$T_c$>, which provides information about the average of all the time intervals between visible events. Applying this concept to the 2003--2012 earthquake sequence in Kachchh, western India, \citet{Telesca2016} found that the variation of <$T_c$> in time exhibited a plausible relationship with the occurrence of earthquakes. In particular, they observed that the value of <$T_c$> decreases significantly before the largest shock in the Kachchh catalog.

In this article we use the visibility graph method to analyze the seismicity of northern Iran from 2005 to 2015. We focus our analysis on these years window because in a previous study we observed the latest decade of seismic records in the region offered the most complete catalog of events \citep[e.g.][]{Khoshnevis2016}. Upon a brief review of the basic concepts of the method and its applications, we describe the seismicity of northern Iran considering three dominant seismic areas, namely: Azerbaijan, Alborz, and Kopeh Dagh. These are regional areas of considerable ground shaking activity, with average seismic rates varying between about 0.17 and 3.25 events per year \citep[e.g.,][]{Nemati2015}. The considerable amount of earthquakes registered in a 10-year period facilitates the analysis done with the visibility graph approach within what would otherwise be considered a relatively short period of time, and serves as a good case study to test the capabilities of the method. We first summarize the dataset, including a description of the seismic catalog and its completeness based on a recent work on the seismicity of the region \citep{Khoshnevis2016}. We then present results obtained for the region regarding the relationship between the the graph's $k$-$M$ slope and the seismicity $b$-values for the three seismic areas, analyze the sensitivity of the catalogs to the number of events and the threshold magnitude, and present results for the variation of <$T_c$> through time.

% NAEEM'S INTRODUCTION
% --------------------

% This article presents the application of VG method on seismicity of northern Iran using the recorded earthquake during the last 10 years (2005-15). Recent studies show that this period (2005-15) has the most complete catalog especially for small earthquakes $(M < 3)$ \citep[e.g.][]{Khoshnevis2016}.  \citet{Lacasa2008} introduced a simple and fast computational method, known as visibility algorithm, to convert time series into graphs. Basically, the visibility graph presents the connection between nodes in a network. This concept has applications in many fields including geophysical studies\citep[e.g.,][]{Telesca2012_CSF,Long2013,Wang2012}.

% The analysis of seismic sequence through visibility graph for various tectonic seismic regions revealed an alternative approach to studying the magnitude time series. \citet{Telesca2012} studied the seismic sequence of Italy between the years 2005 and 2010 using the visibility graph method. Applying different threshold magnitudes and observing the collapsing effect of all distribution degrees, they argued that in analyzing the magnitude time series, VG seems to depend on only the values of the magnitude and being independent of the threshold magnitude.

% In other research, \citet{Telesca2013} studied the seismicity of the Mexican subduction zone through the visibility graph approach. They extracted characteristics of the visibility graph for five different tectonic seismic regions.  According to their study, a visibility graph could distinguish the region with different seismicity characteristics, resulting from its different tectonic processes governing the area. Their study investigated the relationship between the Gutenberg-Richter parameters and the $k-M$ slope, which is the slope of the regression line fitting the magnitude of each event ($M$) and its connectivity degree ($k$).

% In a similar article, \citet{Telesca2014} studied the seismicity of the 2002-11 Pannonian seismicity using visibility graph. They classified the seismic catalog into shallow and deep earthquakes, and extracted the visibility graph characteristics for each class. According to their study, a close relationship exists between the Gutenberg-Richter $b-value$ and the $k-M$ slope of the visibility graph. The high linear correlation coefficient value (close to 1) of $b-value$ and $k-M$  slope suggests the universal character of the relationship between these parameters \citep{Telesca2014}.  Using the synthetic seismicity generated by a simple stick-slip system with asperities, \citet{Telesca2014-pone} investigated the relationship between the $b-value$ of the Gutenberg-Richter law and the slope of the $k-M$ plot as obtained by the VG method. They build up a frictional system as a proxy of geological faults under tectonic stress and defined different asperities by means of modifying the sandpaper grade. They collected data after the first run ($R_1$) and the fifth run ($R_5$) mimicking the younger and mature faults, respectively. They calculated the magnitude as a normalized displacement in range of $1-10$. According to their study the $b-value$ and the $k-M$ slope are linearly correlated to each other with a high correlation coefficient.

% VG method also is applied through considering the timespan between earthquakes. \citet{Telesca2016} defined $<T_c>$ as the parameter window mean interval connectivity time, as an indication of the mean linkage time between earthquakes, in studying the 2003-12 earthquake sequence in the Kachchh region of western India. They found that $<T_C>$ changes through time, indicating that the topological properties of the earthquake network are not stationary. In addition, $<T_c>$ significantly decrees before the largest shock of the catalog. Analyzing the Gutenberg-Richter values of  the magnitude time series of an earthquake sequence revealed more opportunities to study the magnitude time series by using the features of networks.

% In this study, we use visibility graph analysis to investigate the earthquake sequence of northern Iran. We divided northern Iran into three tectonic seismic regions, namely Azerbaijan (Az), Alborz Mountain Range (Al), and Kopeh Dagh (Ko).  Recently, northern Iran has been studied in detail from different seismological points of view.  \citet{Nemati2015} studied the most recent 200 years of seismicity in northern Iran. The frequency of shocks varied widely, from one main shock per 6 years (0.17 event/ year) for the Azerbaijan region to 13 earthquakes per 4 years (3.25 event/year) for the Kopeh-Dagh region \citep{Nemati2015}. Using the smoothed seismicity method, \citet{Khoshnevis2016} studied the seismic hazard of the three northern tectonic seismic regions. Having a considerable number of recorded earthquakes, extensive seismic history, and being in a unique tectonic seismic region, northern Iran is a good platform from which to analyze the visibility graph method's capabilities for studying the magnitude time series. Significant aspect of our work includes the compassion of earthquake catalog, conversion of the earthquake magnitude and computation of the completeness magnitude for each tectonic seismic zone. We present the relationship between  $k$-$M$  slope and $b$-value in these three tectonic seismic regions. We analyze the sensitivity of the catalogs to the number of events and the threshold magnitude. We also investigate the variation of  $<T_c>$  value through time.

>>>>

For the particular case of seismological applications, previous studies have proposed to consider the nodes in the graph to be the earthquakes in a given time series of events, and analyzed the properties of the graph with respect to traditional seismicity parameters. The analysis of seismic sequences through the visibility graph approach for various tectonic seismic regions has proven to be a valid alternative to studying magnitude time series. \citet{Telesca2012}, for instance, studied the seismicity of Italy between the years 2005 and 2010 using the visibility graph method. Applying different threshold magnitudes to construct the graph and observing the collapsing effect of all distribution degrees, they observed that the properties of the visibility graph seemed to depend only on the magnitude values, and not on the threshold magnitude used in the analysis.

@@

In a similar article, \citet{Telesca2014} studied the seismicity of the 2002--2011 Pannonian region using the visibility graph method for two sub-catalogs of shallow and deep earthquakes. They extracted the visibility graph characteristics for each group of events and confirmed that there was a close correlation between the Gutenberg-Richter $b$-value and the slop of the $k$-$M$ relationship obtained from the visibility graph. According to \citet{Telesca2014}, the high linear correlation coefficient value (close to 1.0) between the $b$-value and the $k$-$M$ slope indicates that this relationship exhibits a nearly universal character. This observation was reinforced experimentally by \citet{Telesca2014-pone}, who investigated the behavior of a mechanical stick-slip system with different asperities (sandpaper of different grades). After collecting data for different experiments emulating young and mature faults, \citet{Telesca2014-pone} found again that the $b$-value was linearly correlated with the $k$-$M$ slope derived from the visibility graph of the experimental events.

More recently, \citet{Telesca2016} used the visibility graph method to gain insight about the timespan between earthquakes. The authors defined the parameter <$T_c$> as the window mean interval connectivity time, which provides information about the average of all the time intervals between visible events. Applying this concept to the 2003--2012 earthquake sequence in Kachchh, western India, \citet{Telesca2016} found that the variation of <$T_c$> in time exhibited a plausible relationship with the occurrence of earthquakes. In particular, they observed that the value of <$T_c$> decreases significantly before the largest shock in the Kachchh catalog.

In this article we use the visibility graph method to analyze the seismicity of northern Iran from 2005 to 2015. We focus our analysis on these years because in a previous study we observed that the latest decade of seismic records in the region offered the most complete catalog of events \citep[e.g.][]{Khoshnevis2016}. Upon a brief review of the basic concepts of the method and its applications, we describe the seismicity of northern Iran considering three dominant seismic regions: Azerbaijan, Alborz, and Kopeh Dagh. These are regional areas of considerable ground shaking activity, with average seismic rates varying between about 0.17 and 3.25 events per year \citep[e.g.,][]{Nemati2015}. The considerable amount of earthquakes registered in a 10-year period facilitates the analysis done with the visibility graph approach within what would otherwise be considered a relatively short period of time, and serves as a good case study to test the capabilities of the method. We first summarize the dataset, including a description of the seismic catalog and its completeness based on a recent work on the seismicity of the region \citep{Khoshnevis2016}. We then present results obtained for the region regarding the relationship between the the graph's $k$-$M$ slope and the seismicity $b$-values for the three seismic areas, analyze the sensitivity of the catalogs to the number of events and the threshold magnitude, and present results for the variation of <$T_c$> through time.

### 2016-08-22 methodology.tex
working on last versions of figures in and on results section
fixed minor details in text

Discrete mathematics and computer science often deal with different types of data abstractions and structures such as sets, trees, and graphs, where concepts like nodes and links are used to describe the topology a collection of objects and manipulate their data \citep[e.g.,][]{Skiena_2008_Book}. In graphs, the data---sometimes also referred to as the payload---resides at the vertices or nodes who are connected through links or edges, which may or may not have an associated direction. A visibility graph is a special type of undirected graphs in which the links are straight lines connecting intervisible nodes; that is, straight lines that do not overlap any obstacle while connecting nodes that can see each other in a physical space \citet{LozanoPerez_1979_CACM}. Visibility graphs have been mostly used in robotics for navigation path planning \citep[e.g.,][]{Huang_2004_Proc, Oommen_1987_JRA} but have also seen applications in other fields including urban studies, interior architecture, medicine and geosciences \citep[e.g.,][]{Raman_2010_UE, Ahmadlou_2010_JNT, Varoudis_2014_JSS, Phillips_2015_ESR}.

In a relatively recent study, \citet{Lacasa2008} applied the concepts of visibility graphs to the representation and analysis of time series. Multiple applications have been found to this idea in fields like economics \citep{Yang_2009_PA, Wang2012} and climate \citep{Elsner_2009_GRL}. In this study, we are particularly interested in the application of visibility graphs to the analysis of seismic sequences \citep{Telesca2012}. In this case, the nodes in the visibility graphs are considered to be seismic events distributed over time. For any seismic time series, two characteristics are attributed to each event: (a) its occurrence time, and (b) the value associated to the event, here considered as the magnitude. The obstacles in the time space are the vertical lines (or sticks) between the time axis and the magnitude of the event in a time-magnitude series. Two events are in connection, or visible to each other, whenever no other event interrupts their linear connection.

In mathematical form, events $i$ and $j$ are visible to each other if they satisfy the inequality
%
\begin{equation}
	\frac{y_i - y_p }{t_p - t_i} > \frac{y_i - y_j}{ t_j - t_i} \, ,
\end{equation}
%
\noindent
where $y$ is the value associated with the event and $t$ is the time of the event. The index $p$ indicates any event occurring between events $i$  and $j$. It follows from this that the visibility graph generated from a time series holds the following conditions. Connectivity: each event is visible to the immediate first events to its right and left sides, if there is any. Directivity: the graph is considered undirected by definition; that is, the algorithm explaining the connections between events is developed without defining a direction for the links between the events. And invariance: scaling or time-shifting the series does not change the resulting visibility graph, provided the transformation is done under affine conditions \citep{Lacasa2008}.

Figure \ref{fig:vg} shows an example of a visibility graph for one of the magnitude-time series to be considered later in this study. For each event $i$ we compute the connectivity degree, $k$, which is the sum of the connections to all other events $j$ visible by the $i$th event. Events are often categorized in small magnitude bins with $\Delta M = 0.1$ and we plot the connectivity degree as a function of the bin magnitude. Note that two earthquakes with the same magnitude can have different $k$ values, which ultimately depend on the occurrence time of the event and whether the time-neighboring events are of a larger or smaller magnitude.

@@

Plotting $k$ against magnitude results on a scattered set of points which have been shown to be acceptably represented by a linear correlation. We referred to this correlation as the $k$-$M$ relationship. More important, the slope of the $k$-$M$ relationship also shows a linear relationship with the seismicity $b$-value of the seismic zone under consideration. This is also true for a universal sample of $k$-$M$ slopes and $b$ values \citep{Telesca2013, Telesca2014}. \citet{Telesca2014} also observed that it was reasonable to draw a relationship between the distribution of the events when considering whether these were connected (visible to each other) or not, and time. Let $T_c$ be the interval connectivity time, which is nothing but the time difference between two inter-visible events (see Fig.~\ref{fig:vg}); and <$T_c^e$> be the average of all $T_c$ values computed for event $e$; then it is possible to compute the visibility graph mean interval connectivity time <$T_c$>, which is the mean value of all <$T_c^e$> in a given time sequence of events.

Because <$T_c$> can be obtained for any sub-graph within a larger graph, that is, for any time window in a larger magnitude-time series of events, then it is possible to investigate the evolution of <$T_c$> (and that of the $k$-$M$ slope and the $b$-value) for a moving window sliding along the magnitude-time series. When doing so, \citet{Telesca2016} suggested to associate the values of each sub-sequence with the last event in the sequence. Following this approach, \citet{Telesca2016} found that the variation of <$T_c$> over time exhibited a possible relationship with the occurrence of earthquakes. This and the aforementioned relationship between the $k$-$M$ slope and the $b$-value are aspects we explore next for our region and earthquake-sequence of interest.

% Fig.~\ref{fig:vg} shows initial events of KopehDagh magnitude-time series with connectivity degree of each event.

>>>>

Discrete mathematics and computer science often deal with different types of data abstractions and structures such as sets, trees, and graphs, where concepts like nodes and links are used to describe the topology of a collection of objects and manipulate their data \citep[e.g.,][]{Skiena_2008_Book}. In graphs, the data---sometimes also referred to as the payload---resides at the vertices or nodes who are connected through links or edges, which may or may not have an associated direction. A visibility graph is a special type of undirected graphs in which the links are straight lines connecting intervisible nodes; that is, straight lines that do not overlap any obstacle while connecting nodes that can see each other in a physical space \citep{LozanoPerez_1979_CACM}. Visibility graphs have been mostly used in robotics for navigation path planning \citep[e.g.,][]{Huang_2004_Proc, Oommen_1987_JRA} but have also seen applications in other fields including urban studies, interior architecture, medicine and geosciences \citep[e.g.,][]{Raman_2010_UE, Ahmadlou_2010_JNT, Varoudis_2014_JSS, Phillips_2015_ESR}.

In a relatively recent study, \citet{Lacasa2008} applied the concepts of visibility graphs to the representation and analysis of time series. Multiple applications have been found to this idea in fields such as economics \citep{Yang_2009_PA, Wang2012} and climate studies \citep{Elsner_2009_GRL}. In this study, we are particularly interested in the application of visibility graphs to the analysis of seismic sequences \citep{Telesca2012}. In this case, the nodes in the visibility graphs are considered to be seismic events distributed over time. For any seismic time series, two characteristics are attributed to each event: (a) its occurrence time, and (b) the value associated to the event, here considered as the magnitude. The obstacles in the time space are the vertical lines (or sticks) between the time axis and the magnitude of the event in a time-magnitude series. Two events are in connection, or visible to each other, whenever no other event interrupts their linear connection.

In mathematical form, events $i$ and $j$ are visible to each other if they satisfy the inequality
%
\begin{equation}
	\frac{y_i - y_p }{t_p - t_i} > \frac{y_i - y_j}{ t_j - t_i} \, ,
	\label{eq:vg}
\end{equation}
%
\noindent
where $y$ is the value associated with the event and $t$ is the time of the event. The index $p$ indicates any event occurring between events $i$  and $j$. It follows from this that the visibility graph generated from a time series holds the following conditions. Connectivity: each event is visible to the two events to its immediate right and left sides, if there is any. Directivity: the graph is considered undirected by definition; that is, the algorithm explaining the connections between events is developed without defining a direction for the links between the events. And invariance: scaling or time-shifting the series does not change the resulting visibility graph, provided the transformation is done under affine conditions \citep{Lacasa2008}.

Figure \ref{fig:vg} shows an example of a visibility graph for a portion of one of the magnitude-time series to be considered later in this study. For each event $i$ we compute the connectivity degree, $k$, which is the sum of the connections to all other events $j$ visible by the $i$th event. Events are often categorized in small magnitude bins with $\Delta M = 0.1$ and we plot the connectivity degree as a function of the bin magnitude. Note that two earthquakes with the same magnitude can have different $k$ values, which ultimately depend on the occurrence time of the event and whether the time-neighboring events are of a larger or smaller magnitude.

@@

Plotting $k$ against magnitude results on a scattered set of points which have been shown to be acceptably represented by a linear regression. We referred to this correlation as the $k$-$M$ relationship. More important, the slope of the $k$-$M$ relationship also shows a linear regression with the seismicity $b$-value of the seismic zone under consideration. This is also true for a universal sample of $k$-$M$ slopes and $b$ values \citep{Telesca2013, Telesca2014}. \citet{Telesca2014} also observed that it was reasonable to draw a relationship between the distribution of the events when considering whether these were connected (visible to each other) or not, and time. Let $T_c$ be the interval connectivity time, which is nothing but the time difference between two inter-visible events (see Fig.~\ref{fig:vg}); and <$T_c^e$> be the average of all $T_c$ values computed for event $e$; then it is possible to compute the visibility graph mean interval connectivity time <$T_c$>, which is the mean value of all <$T_c^e$> in a given time sequence of events.

Because <$T_c$> can be obtained for any sub-graph within a larger graph, that is, for any time window in a larger magnitude-time series of events, then it is possible to investigate the evolution of <$T_c$> (and that of the $k$-$M$ slope and the $b$-value) for a moving window---with a constant number of events---sliding along the magnitude-time series. When doing so, \citet{Telesca2016} suggested to associate the values of each sub-sequence with the last event in the sub-sequence. Following this approach, \citet{Telesca2016} found that the variation of <$T_c$> over time exhibited a possible relationship with the occurrence of earthquakes. This and the aforementioned relationship between the $k$-$M$ slope and the $b$-value are aspects we explore next for our region and years-interval of interest.

### 2016-08-22 results.tex
fixed minor details in text

We generate the visibility graph for each sub-region of interest using the time-magnitude sequences shown in Fig.~\ref{fig:mag-time}. The number of nodes in each graph is the same as the (declustered) number of events shown in Table \ref{tab:seismicity}, and the links between inter-visible events are established following the condition in equation (\ref{eq:vg}), as in the example shown in Fig.~\ref{fig:vg}. (We do not include a visualization of the complete sequence graphs because the links are so many, it is only practical to visualize the graphs of short sequences). We then collect information about the number of inter-visibility links associated with each event (connectivity degree, $k$) and categorize the events in magnitude bins of size $\Delta M = 0.1$, as mentioned in the Methodology section. Figure \ref{fig:km} shows the scattered distribution of events in the magnitude-connectivity degree plane for each sub-region. The figure shows that, in general, the value of $k$ increases with $M_w$. Figure \ref{fig:km} also shows the results of obtaining linear $k$-$M$ regressions for each dataset and the values of the $k$-$M$ slopes for the three seismic zones.

@@

Next, we examine the relationship between the $b$-values from Table \ref{tab:seismicity} and the $k$-$M$ slope values shown in Fig.~\ref{fig:km}. Figure \ref{fig:regression} shows the scattered results for $k-M$ slope versus $b$-value for the three tectonic seismic regions in northern Iran along with the data-points obtained for analysis of the magnitude-time sequences of the Mexican subduction zone \citep{Telesca2013} and the Pannoninan seismic zone \citep{Telesca2014}, as well as other experimental results \citep{Telesca2014-pone}. Figure \ref{fig:regression} also include the results of different linear regressions between the $b$-value and the $k$-$M$ slope. Each regression shows the addition of new data-points from different studies over time, with the present study adding the latest set of data. Note that the regression improves as new data-points are added, which is indicated by the correlation coefficient $R$, also shown in the figure. The correlation fitting various regional seismic data was previously pointed out by \citet{Telesca2014}.

@@

To further explore the sensitivity of the graph properties to the number of events in each catalog and the value of the minimum magnitude to be considered, we randomly picked a significant number of sub-sequences from within the initial catalog compiled for each region, and repeated the analysis for each sub-sequence. In total, for each region, we extracted 200 new sub-sequences from the initial (pre-declustering) catalog. The number of events in each sub-sequence was varied randomly but chosen to be large enough to represent the seismic characteristics of each region. In particular, the minimum size of each sub-sequence was set to be $n \geq 150$, whereas the maximum size in the sequence was set to be as large as the original (pre-declustering) catalog (see Table \ref{tab:seismicity}). We forced the random sub-sequences to progress positively in time without altering the natural occurrence of events. In other words, we randomly determined the initial event and the sub-sequence size (number of events to be considered), and then picked that number of events following the initial earthquake in the sub-sequence. Next, we determined the value of $M_c$ and $b$ as previously done for the complete catalogs, created the graph for all events with $M \geq M_c$, and extracted the connectivity degree of the events in each sub-sequence.

@@

Figure \ref{fig:random}a shows scattered points for all the individual $k$-$M$ slope and $b$-value pairs for all the sub-sequences that were randomly picked from the initial catalogs of each region in northern Iran. Figures \ref{fig:random}b and \ref{fig:random}c show the variability of the $k$-$M$ slope and $b$-value, respectively. These latter figures show the mean values of each parameter for all the random sub-sequences and the amplitudes of $\pm 1$ standard deviation, as well as the coefficients of variation (in percentage). Figure \ref{fig:random}a also includes the data points corresponding to the complete catalogs (solid symbols); the points corresponding to the mean values of the $k$-$M$ slope and $b$, from Figs.~\ref{fig:random}b and \ref{fig:random}c; and the universal linear regression from equation (\ref{eq:universal.bm}) (thick line), as well as the linear regressions for northern Iran when using the complete catalogs (dashed, thin line) and the random sub-catalogs (continuous, thin line).

The comparison between the mean values of the random sequences analysis versus the complete catalogs presented in Fig.~\ref{fig:random} indicate there exist only a small bias, which is well within the standard deviation of the different value samples. We note that the values of the $k$-$M$ slope are slightly smaller when done with the random sequence analysis; whereas $b$-values seem more stable. The comparisons of the regressions, however, seem to indicate that the analysis of random sub-catalogs leads to a result more in line with the universal results obtained when considering multiple seismic zones. The linear regress of the random sub-catalogs yields the following equation:
% 
\begin{equation}
	b = 0.080 + 0.093 m \, .
	\label{eq:iran.bm}
\end{equation}

Note that equations (\ref{eq:universal.bm}) and (\ref{eq:iran.bm}) have share similar intercepts with the $b$-value axis, and only slightly different slopes, with the analysis for northern Iran leading to the $b$-value to increase more rapidly with $m$ than the universal data regression suggests. The similarity, however, is a positive sign of the stability of the method.

@@

We now investigate the relationship of the parameters obtained through the visibility graph analysis with time. Here, the visibility graph analysis is done by windows of equal number of events moving in time across the catalog sequence. The number of events in the window is kept fixed independently of the time between them, and the results of the visibility graph analysis are associated with the last event in the window. In this case, we are interested in using a small number of events to capture the relevance of each new event as the window moves in time. We tried different numbers and finally chose 20 events as the moving window sequence size. This selection was also convenient given the smaller total number of declustered events in the catalog of Azerbaijan. When we tested larger number of events, we found that we could not properly observe the changes in the different parameters with time. On the other hand, when we used smaller sequence sizes, the values deviated more significantly from the random and complete catalog results obtained previously. At this point it is also of interest to compute the value of <$T_c$> explained int he Methodology section.

Figure \ref{fig:tc} shows the variation of <$T_c$>, the $k$-$M$ slope, and the $b$-value with time for each of the three seismic zones in northern Iran. Note that the behavior of the $k$-$M$ slope and the $b$-value is very similar along time for all three zones. This is consistent with the results discussed above. In the figure we also highlight the occurrence of relevant earthquakes in each seismic zone in the last decade. We note that there seems to be a correlation between the behavior of the $b$-value in time with the occurrence of some of these events. In particular, some of the events in the figure seem to coincide with a drop in the $b$-value.

The decline in $b$-value before large earthquakes has been studied in other regions before \citep[e.g.,][]{Wyss2000, Wyss2006, Schorlemmer2005, Chan2012}. In the context of the visibility graph analysis, \citet{Telesca2016} observed a decrease in <$T_c$> before the large earthquake of the western India earthquake sequence. We recognize, however, that the lack of larger events in our region of interest in the last decade prevents us from drawing a stronger conclusion on this regard.

>>>>

We generate the visibility graph for each sub-region of interest using the time-magnitude sequences shown in Fig.~\ref{fig:mag-time}. The number of nodes in each graph is the same as the (declustered) number of events shown in Table \ref{tab:seismicity}, and the links between inter-visible events are established following the condition in equation (\ref{eq:vg}), as in the example shown in Fig.~\ref{fig:vg}. (We do not include a visualization of the complete sequence graphs because the links are so many, it is only practical to visualize the graphs of short sequences.) We then collect information about the number of inter-visibility links associated with each event (connectivity degree, $k$) and categorize the events in magnitude bins of size $\Delta M = 0.1$, as mentioned in the Methodology section. Figure \ref{fig:km} shows the scattered distribution of events in the magnitude-connectivity degree plane for each sub-region. The figure shows that, in general, the value of $k$ increases with $M_w$. Figure \ref{fig:km} also shows the results of obtaining linear $k$-$M$ regressions for each dataset and the values of the $k$-$M$ slopes for the three seismic zones.

@@

Next, we examine the relationship between the $b$-values from Table \ref{tab:seismicity} and the $k$-$M$ slope values shown in Fig.~\ref{fig:km}. Figure \ref{fig:regression} shows the scattered results for $k$-$M$ slope versus $b$-value for the three tectonic seismic regions in northern Iran along with the data-points obtained for the analysis of the magnitude-time sequences of the Mexican subduction zone \citep{Telesca2013} and the Pannoninan seismic zone \citep{Telesca2014}, as well as other experimental results \citep{Telesca2014-pone}. Figure \ref{fig:regression} also includes the results of different linear regressions between the $b$-value and the $k$-$M$ slope. Each regression reflects the addition of new data-points from different studies. Note that the regression improves as new data-points are added, which is indicated by the correlation coefficient $R$, also included in the figure. The correlation fitting various regional seismic data was previously pointed out by \citet{Telesca2014}.

@@

To further explore the sensitivity of the graph properties to the number of events in each catalog and the value of the minimum magnitude, we randomly picked a significant number of sub-sequences from within the initial catalog compiled for each region, and repeated the analysis for each sub-sequence. In total, for each region, we extracted 200 new sub-sequences from the initial (pre-declustering) catalog. The number of events in each sub-sequence was varied randomly but chosen to be large enough to represent the seismic characteristics of each region. In particular, the minimum size of each sub-sequence was set to be $n \geq 150$, and the maximum size in the sequence was set to be as large as the original catalog (pre-declustering; see Table \ref{tab:seismicity}). We forced the random sub-sequences to progress positively in time without altering the natural occurrence of events. In other words, we randomly determined the initial event and the sub-sequence size (number of events to be considered), and then picked that number of events following the initial earthquake in the sub-sequence. Next, we determined the value of $M_c$ and $b$ as previously done for the complete catalogs, created the graph for all events with $M \geq M_c$, and extracted the connectivity degree of the events in each sub-sequence.

@@

Figure \ref{fig:random}a shows scattered points for all the individual $k$-$M$ slope and $b$-value pairs for all the sub-sequences that were randomly picked from the initial catalogs of each region in northern Iran (small symbols). Figures \ref{fig:random}b and \ref{fig:random}c show the variability of the $k$-$M$ slope and $b$-value, respectively. These latter figures show the mean values of each parameter for all the random sub-sequences and the amplitudes of $\pm 1$ standard deviation, as well as the coefficients of variation (in percentage). Figure \ref{fig:random}a also includes the data points corresponding to the complete catalogs (large solid symbols); the points corresponding to the mean values of the $k$-$M$ slope and $b$, from Figs.~\ref{fig:random}b and \ref{fig:random}c (large hollow symbols); and the universal linear regression from equation (\ref{eq:universal.bm}) (thick line), as well as the linear regressions for northern Iran when using the complete catalogs (dashed thin line) and the random sub-catalogs (continuous thin line).

The comparison between the mean values of the random sequences analysis versus the complete catalogs presented in Fig.~\ref{fig:random} indicates that there exists only a small bias, which is well within the standard deviation of the different value samples. We note that the values of the $k$-$M$ slope are slightly smaller when obtained with the random sequence analysis, whereas the $b$ values seem more stable. The comparisons of the regressions, however, indicate that the analysis of the random sub-catalogs leads to a result more in line with the universal results obtained when considering multiple seismic zones. The linear regression of the random sub-catalogs yields the following equation:
% 
\begin{equation}
	b = 0.080 + 0.093 m \, .
	\label{eq:iran.bm}
\end{equation}

Note that equations (\ref{eq:universal.bm}) and (\ref{eq:iran.bm}) have similar intercepts with the $b$-value axis, and only slightly different slope constants. In this particular case, the analysis for northern Iran leads to a relationship in which the $b$-value increases more rapidly with $m$ than in the case of the regression obtained for the universal data. The similarity between the two equations, however, is a positive sign of the stability of the method.

@@

We now investigate the relationship of the parameters obtained through the visibility graph analysis with time. Here, the visibility graph analysis is done by windows of equal number of events moving in time across the catalog sequence. The number of events in the window is kept fixed independently of the time between them, and the results are associated with the last event in the window. In this case, we are interested in using a small number of events to capture the relevance of each new event as the window moves in time. We tried different numbers and finally chose 20 events as the moving window sequence size. This selection was also convenient given the smaller total number of declustered events in the catalog of Azerbaijan. When we tested larger number of events, we found that we could not properly observe the changes in the different parameters with time. On the other hand, when we used smaller sequence sizes, the values deviated more significantly from the random and complete catalog results obtained previously. At this point it is also of interest to compute the value of <$T_c$> explained in the Methodology section.

Figure \ref{fig:tc} shows the variation of <$T_c$>, the $k$-$M$ slope, and the $b$-value with time for each of the three seismic zones in northern Iran. Note that the behavior of the $k$-$M$ slope and the $b$-value is very similar along time for all three zones. This is consistent with the results presented before. In the figure, and include the time-magnitude sequences. We note that there seems to be a correlation between the behavior of the $b$-value in time with the occurrence of some of these larger events. In particular, some of the events in the figure seem to coincide with a drop in the $b$-value.

The decline of the $b$-value before large earthquakes has been studied in other regions before \citep[e.g.,][]{Wyss2000, Wyss2006, Schorlemmer2005, Chan2012}. In the context of the visibility graph analysis, \citet{Telesca2016} observed a decrease in <$T_c$> before the large earthquake of the western India earthquake sequence. We recognize, however, that the lack of larger ($M>6$) events in our region of interest in the last decade prevents us from drawing a stronger conclusion on this regard.
